Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spanning 500\,Parsecs

### 2021-04-06 cluster_paper/ms.tex
update ms

\newcommand{\kms}{\,km\,s$^{-1}$}
\newcommand{\ms}{\,m\,s$^{-1}$}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across
  position and velocity space.} The core (black) was analyzed by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident
  with the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) is a concatenation of the studies by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
  used less restrictive membership assignment algorithms described in
  Appendix~\ref{app:clustering}.  The field sample is randomly drawn
  from a $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
  low volume density of the halo stars makes it difficult to visually
  distinguish the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening
    correction has been applied.  The core (black) shows a
    main-sequence and turnoff consistent with stars of a fixed age and
    metallicity.  The halo (blue) is similar, though the binary
    sequence is less defined.  The faintest M dwarfs in the core and
    halo are brighter than field stars (gray) with the same color,
    consistent with the cluster M dwarfs not having yet reached the
    main-sequence.  {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are
    overplotted at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25,
    8.5\}$.  The arrow represents the average reddening correction
    applied to the models.  The opacity of the M-dwarf model
    atmospheres is less well-calibrated for $M_\star \lesssim
    0.45M_\odot$ ($\approx$M2V; $Bp-Rp\approx$2.2); these model values
    are shown with dashed lines.
    \label{fig:hr}
  }
\end{figure*}

@@

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $Bp-Rp$ color.  Stars in
the core appear consistent with
having a fixed age and metallicity, and varying mass.  The halo stars
show a similar sequence, though with greater scatter.  One possible
explanation for the additional scatter could be that the halo is more
contaminated by field stars.  For instance, $\approx5$ red
giants in the halo
must be field interlopers, because their isochronally implied
ages would be inconsistent with that of the cluster.

@@

To determine the reddening correction across \cn, we
queried the 2MASS
DUST
service\footnote{
	\url{http://irsa.ipac.caltech.edu/applications/DUST};
	query performed using the \texttt{astrobase.services.dust} module
	\citep{bhatti_astrobase_2018}.
}
and retrieved the extinction parameters at the positions of each
NGC\,2516 member.  The mean and standard
deviation of the $E(B-V)$ values for the \citet{schlegel_maps_1998}
and \citet{schlafly_measuring_2011} maps were consistent, so we
adopted the average from \citet{schlegel_maps_1998}: $E(B-V)_{\rm
S}=0.206\pm0.039$.  \citet{bonifacio_search_2000} found however that
the \citet{schlegel_maps_1998} maps overestimate the reddening values
when the color excess exceeds about 0.10\,mag. Applying the correction
proposed by \citet{bonifacio_search_2000} leads to our adopted
$E(B-V)=0.102\pm0.019$.  Finally, to convert to Gaia colors we used
the calibration of \citet{stassun_TIC8_2019}, namely $E(Bp-Rp)=1.31
E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded $E(Bp-Rp)=0.134\pm0.025$,
which is used in the plots.  To redden the isochrones, we assumed
$R_V=3.1$, and applied the \citet{odonnell_1994} SED-dependent
extinction law star-by-star through the PARSEC interface.

For the metallicity, we considered a range of super and sub-solar
metallicities to fit as much of the locus from $0.5<Bp-Rp<1.5$ as
possible, and settled on the slightly super-solar $[M/H]=0.06$
\citep{cummings_2011_li_iron}.  Sub-solar metallicities led to model
predictions too blue along the main sequence by $\approx$0.1\,mag.
Our adopted metallicity agrees with the spectroscopic metallicity from
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
\citet{bailey_rv_2018} report a slightly sub-solar metallicity for the
cluster, we prefer super-solar based on the goodness of fit from the
photometry.

Overall, the data and models agree for masses above
$\approx$0.45\,$M_\odot$.  Below this mass, the data and models diverge at
colors redder than $Bp-Rp\approx2.2$, in the sense that the model
isochrones have lower luminosities and bluer colors than the
observations.\footnote{The 100 Myr isochrone does overlap well with the
lower main sequence, but it fails to fit the upper main sequence and
red giants.}  The MIST isochrones showed a comparable disagreement
\citep{choi_mesa_2016}.  Proposed explanations for the discrepancy
between the models and observations have include starspots and
incomplete molecular line lists\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
Ultimately, we adopted the PARSEC isochrones because
\citet{chen_improving_2014} implemented an empirical
temperature-opacity calibration, which leads the PARSEC isochrones to
diverge at slightly lower mass than the MIST models.  Regardless, for
purposes of age-dating the cluster we focus on the main-sequence
turn-off (MSTO), since this is where the models have maximal
predictive power.

@@

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Sets
    $\mathcal{A}$ ({\it left}) and $\mathcal{B}$ ({\it right}) undergo successive stages of automated cleaning,
    (see Section~\ref{subsubsec:cluster}).  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
    Stars in the core and the halo of NGC\,2516 overlap with the Pleiades at $(Bp-Rp)_0$$<$1.2.
    From $1.2<(Bp-Rp)_0<1.7$, 
    some NGC\,2516 members have longer rotation periods than Pleaides members.  
    {\it Bottom row}: \cn\
    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
    appear in red (astrometric binaries; 11\% of stars).  Overplotted
    in orange are sources at least 0.3\,mag above an empirically fit
    cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $Rp<16$.  The difference (65 stars)
is due to 35 stars from 
\citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and 30 stars falling on the TESS chip gaps.  The
$Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
the distance of \cn.

@@

The top row of Figure~\ref{fig:rot} shows the resulting rotation
periods, with set $\mathcal{A}$ shown the left and $\mathcal{B}$ on
the right.  While set $\mathcal{A}$ is more complete than set
$\mathcal{B}$, this completeness comes at the expense of greater
contamination.  For instance, stars above the slow sequence (perhaps
field contaminants) mostly disappear in set $\mathcal{B}$.
The ``void'' or ``gap'' underneath the slow sequence is also more
pronounced in set $\mathcal{B}$.  To facilitate a comparison against
the Pleiades and Praesepe, we used the rotation periods and dereddened
colors from Table~5 of \citet{curtis_rup147_2020}, which drew on data
from \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.  The first order conclusion is that the Pleiades and
\cn\ appear gyrochronologically coeval for colors spanning
$0.5<(Bp-Rp)_0<1.2$.  The implications of this comparison for the age
of \cn\ are discussed in Section~\ref{disc:absage}.
%An important note is that stars redder than $(Bp-Rp)_0\gtrsim 2.2$
%(later than $\approx$M2V)  are beyond the magnitude limit of our survey.

The bottom row of Figure~\ref{fig:rot} shows the same data, with the
points colored with indicators of binarity.  To assess astrometric
binarity, we used the renormalized unit weight error (RUWE) from Gaia
EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
1.2 as an astrometric binary (11\% of the overall cluster sample; see
{\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
photometric binarity, we fitted a spline to Figure~\ref{fig:hr} and fixed
the nodes by hand.  We then labeled any points brighter than 0.3\,mag above the cluster
sequence as photometric binaries.  This included 20\% of the overall
cluster sample.  Many of the resulting candidate binaries, though not
all of them, appear below the slow rotation sequence.  In set
$\mathcal{B}$, over the color range where the slow and fast sequence
are present ($0.5<(Bp-Rp)_0<1.3$), there are 281 stars, of which 103
(37\%) show signs of binarity.  Dividing these 281 stars into ``fast''
and ``slow'' sequences by eye,  the fraction of stars showing signs of
binarity in the slow and fast subsamples are 30\% (63/210)
and 56\% (40/71), respectively.  We speculate on possible explanations
in Section~\ref{disc:rotn}.

@@

	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf Members of \cn\ exist hundreds of parsecs from the core, and up
  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
  2D-tangential velocities for halo members that meet kinematic
  membership criteria (Subset~$\mathcal{A}$).  Stars with detected
  rotation periods are shown in blue;  halo members for which rotation
  periods should have been detectable, but were not detected, are
  shown in orange.  Stars in the core are shown as black points.
  Computing the 2D-tangential velocity requires a correction for
  projection effects (Appendix~\ref{app:vproj}).  In the {\it top
  right}, the histograms on bin versus 3D separation from the core and
  2D tangential velocity difference.  Bin widths are chosen to enforce
  the same number of stars in the denominator (54) per bin.  In the
  {\it bottom panels}, the innermost and outermost 100 stars in
  position and velocity space are displayed in the space of rotation
  periods and dereddened colors.
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
comparison, we only show stars with $0.5<(Bp-Rp)_0<1.2$ for which our
TESS pipeline succeeding in making light curves.  In other words---the
base sample is the stars for which we could have plausibly detected a
rotation period.  This color range is preferred since the the slow
sequence becomes less defined for spectral types later than
$\approx$K4V.  The bins in the histograms (top-right) were chosen to
ensure that the same number of stars were in each bin.  Rotation
periods are detected for 66\% of stars at least 25\,pc from the
cluster core.  The average detection rate inside of 25\,pc, 73\%, is
slightly higher.

@@

Speaking in the positive though, our rotation period measurements show
that the halo extends to separations of $\approx$250\,pc in physical
space from the cluster core.  The most widely separated F2V-K2V stars
with rotation periods (blue points in
Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
The total length of the structure is therefore 400--500\,pc, depending
on which members of the halo are chosen as the ``tips'' on either end.
This agrees with the overall structure of the halo reported by
\citetalias{kounkel_untangling_2019}, with some minor caveats visible
in the top-left panels of Figure~\ref{fig:physical_x_rotn}.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		%\subfloat{
			\includegraphics[width=1\textwidth]{f5a.pdf}
		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
  Equivalent width of the 6708\,\AA\ doublet is shown versus
  extinction-corrected color for all kinematic \cn\ members with
  Gaia-ESO ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH
  spectra comprise slightly over half of the halo stars, due to the
  non-targeted selection function of that survey.  The gray points are
  Keck-HIRES measurements of stars with detected transiting planets
  from the Kepler mission \citep{berger_identifying_2018}.  Points
  with ${\rm EW}\approx0\,$m\AA\ are non-detections.
  \label{fig:lithiumcorehalo}
  }
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		%\subfloat{
		\includegraphics[width=1\textwidth]{f5b.pdf}
		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Lithium and rotation in NGC\,2516.}
    Points show all kinematic \cn\ members with TESS rotation periods
    in the Gaia-ESO or GALAH surveys, and are colored by equivalent
    width.  Points with reported equivalent widths below 20\,m\AA\
    (including non-detections) are shown with gray crosses.  The
    lithium-rotation correlation for the K-dwarfs has been observed in
    other young clusters, and is discussed in
    Section~\ref{subsec:lithium}.
		\label{fig:lithiumrot}
	}
\end{figure*}

@@

Gaia-ESO selected candidate \cn\ members to observe with the GIRAFFE
and UVES spectrographs based on previously reported literature members
and publicly available photometry.  Since the existence of the \cn\
halo was not known at the time of target selection, very few ``halo''
stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
ultimately reported stellar parameters (including lithium equivalent
widths and metallicities) for 796 stars that they considered possible
\cn\ members.  Cross-matching against our kinematic list of
\nkinematic\ candidate members by position and imposing a 0$\farcs$5
maximum separation limit yields 492 kinematic members for which a
Gaia-ESO spectrum was acquired and stellar parameters were reported.
15 of these (all with $(Bp-Rp)_0 > 2.0$) are spurious matches based on
the Gaia color and GES effective temperature, and we remove them.
This yields 477 stars, of which 436 are in the core, and 41 are in the
halo.  The lop-sided ratio is due to the Gaia-ESO selection function.
We verified by directly querying the ESO archive that
\citet{randich_gaiaeso_2018} indeed included $>99\%$ of the available
Gaia-ESO spectra as part of their \cn\ sample, and so we opted to use
their results rather than repeating their analysis.

@@

Figure~\ref{fig:lithiumcorehalo} shows the concatenation of the
Gaia-ESO and GALAH results, with the points colored by whether the
stars are in the core or the halo of the cluster.  The GALAH spectra
span a color range of $0<(Bp-Rp)_0<1.5$, due to the $V<14$ brightness
cutoff of the survey.  The overall increase of Li EW from
$0<(Bp-Rp)_0<1.0$ is thought to be driven by the temperatures in the
stellar chromospheres: hotter stars fully ionize Li out of its ground
state \citep[{\it e.g.}, Figure~4 of][]{soderblom_evolution_1993}.  The
depletion of stars redder than $(Bp-Rp)_0\gtrsim1.4$, {\it i.e.},
later than K4.5V (0.71\,$M_\odot$), is visible relative to the earlier
K dwarfs.  For comparison, we have also plotted the Li EWs measured by
\citet{berger_identifying_2018} for planet-hosting stars in the Kepler
field (we interpolated from their effective temperatures using the
\citet{pecaut_mamajek_2013}
table\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version 2019.3.22.}.  The field star distribution peaks at late F-type
stars.  The majority of kinematically selected stars in the core and
the halo show lithium equivalent widths substantially in excess of
these field stars.

@@

\subsection{Contamination fraction across the halo}

The main aim this study is to verify whether stars
in the halo of \cn\ are the same age as stars in the core, and
secondarily to quantify the contamination rate.
Figure~\ref{fig:physical_x_rotn}

@@

% \subsubsection{The slowest K dwarfs of NGC\,2516}
% %FIXME need to rewrite this paragraph? (Or maybe not, because it's
% %already in the absolute age section...
% A few conclusions can be drawn by comparing the \cn\ rotation periods
% with those of the Pleiades and Praesepe.  The first is that in both
% stages of cleaning, the core and halo of \cn\ each show
% gyrochronological ``slow sequences''.  These sequences appear to be
% coeval with each other, and at first order appear to overlap
% with the Pleiades sequence.
% 
%   \paragraph{Stalling}
%   Curtis19 found NGC6811 was stalling in the K dwarfs, i.e., Skumanich
%   spindown is not monotonic in time.
%   We're seeing latest K dwarfs spin slowest, earlier FGKs don't show
%   much change.

%   \paragraph{Are the ``very slow rotators'' bonafide members?}
%   Unsure. They are not isotropically distributed around the
%   cluster... so either triggered star formation (CITE, CITE), or they're
%   actually field stars.

>>>>

\newcommand{\kms}{\,km\,s$^{-1}$}
\newcommand{\ms}{\,m\,s$^{-1}$}
\newcommand{\bpmrp}{(\mathrm{Bp}-\mathrm{Rp})_0}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across
  position and velocity space.} The core (black) was analyzed by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident
  with the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) is a concatenation of studies by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
  used less restrictive membership assignment algorithms described in
  Appendix~\ref{app:clustering}.  The field sample is randomly drawn
  from a $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
  low volume density of the halo stars makes it difficult to visually
  distinguish the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening
    correction has been applied.  The core (black) shows a
    main-sequence and turnoff consistent with stars of a fixed age and
    metallicity.  The halo (blue) is similar, though the binary
    sequence is less defined.  The faintest M dwarfs in the core and
    halo are brighter than field stars (gray) with the same color,
    consistent with the cluster M dwarfs not having yet reached the
    main-sequence.  {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are
    overplotted at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25,
    8.5\}$.  The arrow represents the average reddening correction
    applied to the models.  The opacity of the M-dwarf model
    atmospheres is less well-calibrated for $M_\star \lesssim
    0.45M_\odot$ ($\approx$M2V; $\mathrm{Bp}-\mathrm{Rp}\approx$2.2); these model values
    are shown with dashed lines.
    \label{fig:hr}
  }
\end{figure*}

@@

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $\mathrm{Bp}-\mathrm{Rp}$ color.  Stars in
the core appear consistent with
having a fixed age and metallicity, and varying mass.  The halo stars
show a similar sequence, though with greater scatter.  One possible
explanation for the additional scatter could be that the halo is more
contaminated by field stars.  For instance, $\approx5$ red
giants in the halo
must be field interlopers, because their isochronally implied
ages would be inconsistent with that of the cluster.

@@

To determine the reddening correction across \cn, we
queried the 2MASS
DUST
service\footnote{
	\url{http://irsa.ipac.caltech.edu/applications/DUST};
	query performed using the \texttt{astrobase.services.dust} module
	\citep{bhatti_astrobase_2018}.
}
and retrieved the extinction parameters at the positions of each
NGC\,2516 member.  The mean and standard
deviation of the $E(\mathrm{B}-\mathrm{V})$ values for the \citet{schlegel_maps_1998}
and \citet{schlafly_measuring_2011} maps were consistent, so we
adopted the average from \citet{schlegel_maps_1998}: $E(\mathrm{B}-\mathrm{V})_{\rm
S}=0.206\pm0.039$.  \citet{bonifacio_search_2000} found however that
the \citet{schlegel_maps_1998} maps overestimate the reddening values
when the color excess exceeds about 0.10\,mag. Applying the correction
proposed by \citet{bonifacio_search_2000} leads to our adopted
$E(\mathrm{B}-\mathrm{V})=0.102\pm0.019$.  Finally, to convert to Gaia colors we used
the calibration of \citet{stassun_TIC8_2019}, namely $E(\mathrm{Bp}-\mathrm{Rp})=1.31
E(\mathrm{B}-\mathrm{V})$ and $A_G=2.72 E(\mathrm{B}-\mathrm{V})$.  This yielded $E(\mathrm{Bp}-\mathrm{Rp})=0.134\pm0.025$,
which is used in the plots.  To redden the isochrones, we assumed
$R_V=3.1$, and applied the \citet{odonnell_1994} SED-dependent
extinction law star-by-star through the PARSEC interface.

For the metallicity, we considered a range of super and sub-solar
metallicities to fit as much of the locus from $0.5<\mathrm{Bp}-\mathrm{Rp}<1.5$ as
possible, and settled on the slightly super-solar $[M/H]=0.06$
\citep{cummings_2011_li_iron}.  Sub-solar metallicities led to model
predictions too blue along the main sequence by $\approx$0.1\,mag.
Our adopted metallicity agrees with the spectroscopic metallicity from
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
\citet{bailey_rv_2018} report a slightly sub-solar metallicity for the
cluster, we prefer super-solar based on the goodness of fit from the
photometry.

Overall, the data and models agree for masses above
$\approx$0.45\,$M_\odot$.  Below this mass, the data and models diverge at
colors redder than $\mathrm{Bp}-\mathrm{Rp}\approx2.2$, in the sense that the model
isochrones have lower luminosities and bluer colors than the
observations.\footnote{The 100 Myr isochrone does overlap well with the
lower main sequence, but it fails to fit the upper main sequence and
red giants.}  The MIST isochrones showed a comparable disagreement
\citep{choi_mesa_2016}.  Proposed explanations for the discrepancy
between the models and observations have include starspots and
incomplete molecular line lists\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
Ultimately, we adopted the PARSEC isochrones because
\citet{chen_improving_2014} implemented an empirical
temperature-opacity calibration, which leads the PARSEC isochrones to
diverge at slightly lower mass than the MIST models.  Regardless, for
purposes of age-dating the cluster we focus on the main-sequence
turn-off (MSTO), since this is where the models have maximal
predictive power.

@@

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Sets
    $\mathcal{A}$ ({\it left}) and $\mathcal{B}$ ({\it right}) undergo successive stages of automated cleaning,
    (see Section~\ref{subsubsec:cluster}).  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
    Stars in the core and the halo of NGC\,2516 overlap with the Pleiades at $\bpmrp$$<$1.2.
    From $1.2<\bpmrp<1.7$, 
    some NGC\,2516 members have longer rotation periods than Pleaides members.  
    {\it Bottom row}: \cn\
    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
    appear in red (astrometric binaries; 11\% of stars).  Overplotted
    in orange are sources at least 0.3\,mag above an empirically fit
    cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $Rp<16$.  The difference (65 stars)
is due to 35 stars from 
\citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and 30 stars falling on the TESS chip gaps.  The
$Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
roughly to $\bpmrp$ of 2.2, or a spectral type of $\approx$M2V, at
the distance of \cn.

@@

The top row of Figure~\ref{fig:rot} shows the resulting rotation
periods, with set $\mathcal{A}$ shown the left and $\mathcal{B}$ on
the right.  While set $\mathcal{A}$ is more complete than set
$\mathcal{B}$, this completeness comes at the expense of greater
contamination.  For instance, stars above the slow sequence (perhaps
field contaminants) mostly disappear in set $\mathcal{B}$.
The ``void'' or ``gap'' underneath the slow sequence is also more
pronounced in set $\mathcal{B}$.  To facilitate a comparison against
the Pleiades and Praesepe, we used the rotation periods and dereddened
colors from Table~5 of \citet{curtis_rup147_2020}, which drew on data
from \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.  The first order conclusion is that the Pleiades and
\cn\ appear gyrochronologically coeval for colors spanning
$0.5<\bpmrp<1.2$.  The implications of this comparison for the age
of \cn\ are discussed in Section~\ref{disc:absage}.
%An important note is that stars redder than $\bpmrp\gtrsim 2.2$
%(later than $\approx$M2V)  are beyond the magnitude limit of our survey.

The bottom row of Figure~\ref{fig:rot} shows the same data, with the
points colored with indicators of binarity.  To assess astrometric
binarity, we used the renormalized unit weight error (RUWE) from Gaia
EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
1.2 as an astrometric binary (11\% of the overall cluster sample; see
{\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
photometric binarity, we fitted a spline to Figure~\ref{fig:hr}, fixing
the nodes by hand.  We then labeled any points brighter than 0.3\,mag above the cluster
sequence as photometric binaries.  This included 20\% of the overall
cluster sample.  Many of the resulting candidate binaries, though not
all of them, appear below the slow rotation sequence.  In set
$\mathcal{B}$, over the color range where the slow and fast sequence
are present ($0.5<\bpmrp<1.3$), there are 281 stars, of which 103
(37\%) show signs of binarity.  Dividing these 281 stars into ``fast''
and ``slow'' sequences by eye,  the fraction of stars showing signs of
binarity in the slow and fast subsamples are 30\% (63/210)
and 56\% (40/71), respectively.  We speculate on possible explanations
in Section~\ref{disc:rotn}.

@@

	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf Members of \cn\ exist hundreds of parsecs from the core, and up
  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
  2D-tangential velocities for halo members that meet kinematic
  membership criteria (Subset~$\mathcal{A}$).  Stars with detected
  rotation periods are shown in blue;  halo members for which rotation
  periods should have been detectable, but were not detected, are
  shown in orange.  Stars in the core are shown as black points.
  Computing the 2D-tangential velocity requires a correction for
  projection effects (Appendix~\ref{app:vproj}).  In the {\it top
  right}, the histograms on bin versus 3D separation from the core and
  2D tangential velocity difference.  Bin widths are chosen to enforce
  the same number of stars in the denominator ($N=54$) per bin.  In the
  {\it bottom panels}, the innermost and outermost 100 stars in
  position and velocity space are displayed in the space of rotation
  periods and dereddened colors.
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
comparison, we only show stars with $0.5<\bpmrp<1.2$ for which our
TESS pipeline succeeding in making light curves.  In other words---the
base sample is the stars for which we could have plausibly detected a
rotation period.  This color range is preferred since the the slow
sequence becomes less defined for spectral types later than
$\approx$K4V.  The bins in the histograms (top-right) were chosen to
ensure that the same number of stars were in each bin.  Rotation
periods are detected for 66\% of stars at least 25\,pc from the
cluster core.  The average detection rate inside of 25\,pc, 73\%, is
slightly higher.

@@

Even with contaminants, our rotation period measurements show
that the halo extends to separations of $\approx$250\,pc in physical
space from the cluster core.  The most widely separated F2V-K2V stars
with rotation periods (blue points in
Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
The total length of the structure is therefore 400--500\,pc, depending
on which members of the halo are chosen as the ``tips'' on either end.
This agrees with the overall structure of the halo reported by
\citetalias{kounkel_untangling_2019}, with some minor caveats visible
in the top-left panels of Figure~\ref{fig:physical_x_rotn}.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		%\subfloat{
			\includegraphics[width=1\textwidth]{f5a.pdf}
		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
  Equivalent width of the 6708\,\AA\ doublet is shown versus
  dereddened color for all kinematic \cn\ members with
  Gaia-ESO ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH
  spectra comprise slightly over half of the halo stars, due to the
  non-targeted selection function of that survey.  The gray points are
  Keck-HIRES measurements of stars with detected transiting planets
  from the Kepler mission \citep{berger_identifying_2018}.  Points
  with ${\rm EW}\approx0\,$m\AA\ are non-detections.
  \label{fig:lithiumcorehalo}
  }
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		%\subfloat{
		\includegraphics[width=1\textwidth]{f5b.pdf}
		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Lithium and rotation in NGC\,2516.}
    Points show all kinematic \cn\ members with TESS rotation periods
    in the Gaia-ESO or GALAH surveys, and are colored by equivalent
    width.  Points with reported equivalent widths below 20\,m\AA\
    (including non-detections) are shown with gray crosses.  The
    lithium-rotation correlation for the K-dwarfs has been observed in
    other young clusters, and is discussed in
    Section~\ref{subsec:lithium}.
		\label{fig:lithiumrot}
	}
\end{figure*}

@@

Gaia-ESO selected candidate \cn\ members to observe with the GIRAFFE
and UVES spectrographs based on previously reported literature members
and publicly available photometry.  Since the existence of the \cn\
halo was not known at the time of target selection, very few ``halo''
stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
ultimately reported stellar parameters (including lithium equivalent
widths and metallicities) for 796 stars that they considered possible
\cn\ members.  Cross-matching against our kinematic list of
\nkinematic\ candidate members by position and imposing a 0$\farcs$5
maximum separation limit yields 492 kinematic members for which a
Gaia-ESO spectrum was acquired and stellar parameters were reported.
15 of these (all with $\bpmrp > 2.0$) are spurious matches based on
the Gaia color and GES effective temperature, and we remove them.
This yields 477 stars, of which 436 are in the core, and 41 are in the
halo.  The lop-sided ratio is due to the Gaia-ESO selection function.
We verified by directly querying the ESO archive that
\citet{randich_gaiaeso_2018} indeed included $>99\%$ of the available
Gaia-ESO spectra as part of their \cn\ sample, and so we opted to use
their results rather than repeating their analysis.

@@

Figure~\ref{fig:lithiumcorehalo} shows the concatenation of the
Gaia-ESO and GALAH results, with the points colored by whether the
stars are in the core or the halo of the cluster.  The GALAH spectra
span a color range of $0<\bpmrp<1.5$, due to the $V<14$ brightness
cutoff of the survey.  The overall increase of Li EW from
$0<\bpmrp<1.0$ is thought to be driven by the temperatures in the
stellar chromospheres: hotter stars fully ionize Li out of its ground
state \citep[{\it e.g.}, Figure~4 of][]{soderblom_evolution_1993}.  The
depletion of stars redder than $\bpmrp\gtrsim1.4$, {\it i.e.},
later than K4.5V (0.71\,$M_\odot$), is visible relative to the earlier
K dwarfs.  For comparison, we have also plotted the Li EWs measured by
\citet{berger_identifying_2018} for planet-hosting stars in the Kepler
field (we interpolated from their effective temperatures using the
\citet{pecaut_mamajek_2013}
table\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version 2019.3.22.}.  The field star distribution peaks at late F-type
stars.  The majority of kinematically selected stars in the core and
the halo show lithium equivalent widths substantially in excess of
these field stars.

@@

% \subsection{Contamination fraction across the halo}
% 
% The main aim this study is to verify whether stars
% in the halo of \cn\ are the same age as stars in the core, and
% secondarily to quantify the contamination rate.
% Figure~\ref{fig:physical_x_rotn}

@@
